Differentiable transit models for ellipsoidal planets

### 2024-05-23 src/tex/ms.tex
[showyourwork] overleaf sync

% Main body with filler text
\section{Introduction}
The contemporary study of exoplanets frequently assumes spherical bodies. While this is often a valid assumption, the deviations from this case are informative and contain valuable information which may be discarded by assuming sphericity.
\label{sec:intro}

@@

\twocolumngrid
We note that other methods of finding the flux in transit of elliptical occultors also require solving for quartic polynomials \citep{rein2023}. We solve the roots of this polynomial using eigendecomposition of the companion matrix (see \ref{sec:jax0planet} for details of the implementation). \textcolor{red}{we need to update this in eclipsoid to use a more numerically precise method}. The solution gives the x values of the intersection points where the projected disk of the star and the occultor coincide. We can then find the corresponding y values by plugging it back into the formula for either the occultor or occulted body.

@@

\subsection{Limb darkening}
\subsection{Oblate Case}
\subsection{Prolate Case}
\section{Implementation in \lowercase{\texttt{jax0planet}}} \label{sec:jax0planet}
-implementation details:\\
    -API\\
    -root finding (with custom derivatives)\\
    -gauss quad\\
    -plot showing it matches brute force for each term in Gn   \\
\begin{figure}[ht!]
    \script{bruteforce_comparison.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/bruteforce_comparison.pdf}
        \caption{Comparison of the 1D Green's integral to a brute force 2D integral for each term in Green's basis. Residuals are under the expected numerical error in the brute force solution for all terms.
        }
        \label{fig:bruteforce_comparison}
    \end{centering}
\end{figure}

>>>>

% Main body with filler text
\section{Introduction} \label{sec:intro}
The contemporary study of exoplanets frequently assumes spherical bodies. While this is often a valid assumption, the deviations from this case are informative and contain valuable information which may be discarded by assuming sphericity.

In Sec.~\ref{sec:model}, we describe a new forward model for light curves of spherical bodies with ellipsoidal occultors. We then describe the implementation of this model in the Python package \texttt{eclipsoid} in Sec.~\ref{sec:eclipsoid}. The package is built using the \texttt{jax} framework, which allows automatic differentiation of the model with respect to its parameters. We describe a use of this in Sec.~\ref{sec:jwstdetect}, where we utilize Fisher information to quantify the detectability of ellipsoidal exoplanets using JWST and other instruments. Lastly, in Sec.~\ref{sec:lctest}, we apply the new model to the JWST white light curve of WASP107b and demonstrate that we can constrain the oblateness and prolateless of the exoplanet.

This paper has been written using the open-source package \textcolor{red}{\textit{showyourwork!}} for transparency and ease of reproduction. At the end of every figure caption, we link to a GitHub repository containing a Python script used to generate the figure. All data used in this paper is hosted on Zenodo at the following link.

@@

\twocolumngrid
We note that other methods of finding the flux in transit of elliptical occultors also require solving for quartic polynomials \citep{rein2023}. We solve the roots of this polynomial using eigendecomposition of the companion matrix (see \ref{sec:eclipsoid} for details of the implementation). \textcolor{red}{we need to update this in eclipsoid to use a more numerically precise method}. The solution gives the x values of the intersection points where the projected disk of the star and the occultor coincide. We can then find the corresponding y values by plugging it back into the formula for either the occultor or occulted body.

@@

\subsection{Limb darkening}
\subsection{Oblate Case}
\subsection{Prolate Case}
\section{Implementation in eclipsoid} \label{sec:eclipsoid}
-implementation details:\\
    -API\\
    -root finding (with custom derivatives)\\
    -gauss quad\\
    -plot showing it matches brute force for each term in Gn   \\
\begin{figure}[ht!]
    \script{bruteforce_comparison.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/bruteforce_comparison.pdf}
        \caption{Comparison of the 1D Green's integral to a brute force 2D integral for each term in Green's basis. Residuals are under the expected numerical error in the brute force solution for all terms.
        }
        \label{fig:bruteforce_comparison}
    \end{centering}
\end{figure}
